\color{navyblue} Difference-in-Differences Estimation with Spatial Spillover

### 2021-02-14 latex/paper/Spillover.tex
Rework simulation section

The rest of the paper is structured as follows. Section \ref{sec:po_framework} presents the potential outcomes framework, defines the estimand of interest, and shows the resulting bias from estimating a classical difference-in-differences model. Section \ref{sec:monte_carlo} presents Monte Carlo simulations to illlustrate the bias result and to evaluate currently used solutions in the literature. Section \ref{sec:tva} presents an application for evaluating the Tennessee Valley Authority program. Section \ref{sec:event_study} discusses briefly how to incorporate spillovers into event-study estimation following the insights from \citet{Callaway_Sant’Anna_2020}.

@@

\begin{example}
    Last, is the spatial decay function where exposure decreases with distance. In this case, spillover intensity is the sum across all treated observations' decay term, i.e. \begin{equation}\label{eq:h_decay}
        h(\vec{D}, i) = \sum_{j \neq i} D_j e^{-\alpha d(i,j)}.
    \end{equation} 
    This exposure mapping allows for the intensity of spillovers to depend on distance to treatment and also is additive in the number of nearby units treated.\footnote{However, this specification assumes that all units are affected by all other units. This creates problems with inference because it implies potential correlation between all units' error terms. For this reason, this function often is summed over only the $k$-nearest neighbors or over units within $\bar{d}$ miles.} The speed of spillover decay over distance depends on the parameter $\alpha$ that can be calibrated by the researcher or estimated using non-linear least squares. In the literature on R\&D investment, \citet{Keller_2002} uses a modified version of this exposure mapping where $D_j$ is country $j$'s R\&D expenditure. 
\end{example}

@@

\begin{proposition}[Bias from Difference-in-Differences Estimation]\label{thm:bias}\ \\    
    If Assumption \ref{eq:parallel} holds, the expectation of the estimate $\hat{\tau}$ from (\ref{eq:twfe}) is
    \begin{align*}
        \mathbb{E}[\hat{\tau}] &= \underbrace{\mathbb{E}\left[ Y_{i1} - Y_{i0} \ \vert \ D_i = 1 \right] - \mathbb{E}\left[ Y_{i1} - Y_{i0} \ \vert \ D_i = 0 \right]}_{\text{Difference-in-Differences}} \\ 
        &= 
        \mathbb{E} \left[ Y_{i1}(1, \vec{0}) - Y_{i1}(0, \vec{0}) \mid D_i = 1 \right] + \mathbb{E} \left[ Y_{i1}(1, h(\vec{D}, i)) - Y_{i1}(1, \vec{0}) \mid D_i = 1 \right] \\
        &\quad - \mathbb{E} \left[ Y_{i1}(0, h(\vec{D}, i)) - Y_{i1}(0, \vec{0}) \mid D_i = 0 \right] \\
        &= \tau_{\text{direct}} + \tau_{\text{spill,treated}} - \tau_{\text{spill,control}}
    \end{align*}
\end{proposition}

@@

% ------------------------------------------------------------------------------
\subsection{Improving Estimation by Parameterizing Spillovers}
% ------------------------------------------------------------------------------

The previous section shows how to estimate the magnitude of potential bias in the estimation of (\ref{eq:twfe}). However, parameterization of the potential outcomes and exposure mapping allows for unbiased estimates of $\tau_{\text{direct}}$ by controlling for the exposure mapping directly. This is recommended for two reasons.

First, controlling for exposure mapping allows for unbiased estimates of the direct effect of treatment if the potential outcomes are correctly specified. In the case where potential outcomes are `somewhat' incorrectly specified, this can remove some of the bias in estimation. Simulations below highlight that since spillovers are likely to decay over distance, controlling for immediate neighbors removes a large portion of the average spillover effects even under misspecification.

Second, the spillover effects themselves are potentially relevant for policy makers. If treatment for a given area is positive and there are positive benefits to neighboring regions, the estimated benefits will double-understate the true effect of treatment. First, it will not include the neighboring units' benefit in the calculation. Second, since the treatment effect will subtract the neighboring units' benefit, the neighboring units' benefit will be \emph{double undercounted}. Similarly, a policy maker trying to determine the net benefits for a treated location must consider the direct effect and the spillover effect onto treated units. This is because the size of benefits created by the treatment depend on the number of treated units nearby.

@@

Second, a researcher must decide how far the spillovers can reasonably be experience and whether the effect decays over distance. Things such as driving distance or access via public transportation can be helpful in deciding which units are `close enough' to experience spillovers. There are two commonly used techniques to model treatment effect decay.\footnote{If the spillovers is due to crossing borders into the treated area, then a reasonable decision about whether the effect decays is the frequency in which people use the good. For example, libraries whose benefit is only experienced from frequent usage would suggest that the effect decays over distance. On the other hand, abortion clinics opening are used infrequently and therefore the effect likely doesn't decay over distance within a certain cutoff.} First, as seen above is to use an exponential decay function with regards to distance. The other option is to use a set of `concentric donut' indicators which are easy to interpret. An example of this is shown in the empirical application in Section \ref{sec:tva}.

Last, it is important for estimation to know if the spillovers are additive or non-additive in nature. If the number of nearby treated units matter, then the exposure mapping should sum over nearby treated units in some way. For example, in the context of library construction, the number of neighbors with libraries does not change the size of the spillover for it is access to \textit{any} library that matters. In the context of factory openings, though, agglomeration economies imply that the spillovers are additive in nature.

% ------------------------------------------------------------------------------
\section{Monte Carlo Simulations}
\label{sec:monte_carlo}
% ------------------------------------------------------------------------------

% ------------------------------------------------------------------------------
\subsection{Spillover on Controls Only}
% ------------------------------------------------------------------------------

I now turn to a series of Monte Carlo Simulations to highlight the importance for controlling for spillovers in simulations. I use the set of counties in the contiguous United States and the data generating process used is defined by the exposure mapping given in (\ref{eq:example_exposure}) with cutoff distance $\bar{d} = 40$ miles and the data generating process given in (\ref{eq:example_po}). A unit of observation is a US county and the periods are $t \in \{1, \dots, 20\}$ with treatment turning on after period $t = 10$. The unit and time fixed effects are generated by $\mu_t \sim N(0.2t, 0.1^2)$ and $\mu_i \sim N(6, 2^2)$ respectively, and the error term is $\varepsilon \sim N(0, 2^2)$. Last, the size of treatment and spillovers are as follows: $\beta_{\text{direct}} = 2$, $\beta_{\text{spill, control}} = 1$ and $\beta_{\text{spill, treat}} = 0$.

For the first simulation, I restrict spillovers only to occur among control units, i.e. $\beta_{\text{spill, treat}} = 0$, and estimate the two-way fixed effects model, (\ref{eq:twfe}). This data generating process where spillovers only occur onto control units matches what has been typically assumed in the literature. I assign treatment among counties randomly with various unconditional probabilities between 3 percent and 50 percent. The data-generating process is therefore 
\begin{equation}
    \label{eq:dgp1} 
    y_{it} = \mu_t + \mu_i + \beta_{\text{direct}} D_{it} + \beta_{\text{spill, control}} (1-D_{it}) \text{Near}_{it} + \varepsilon_{it}   
\end{equation}

% -----------------------------------------------------------------------------
\subsection{Spillovers onto Control and Treated Units}
% ------------------------------------------------------------------------------

% ------------------------------------------------------------------------------
\subsection{Misspecification of Spillovers}
% ------------------------------------------------------------------------------

Now we turn to simulations where we include potentially incorrectly specified exposure mappings and see their performance in a set of different data generating processes. In the general form, I generate data with the same fixed effects and error term as before, but for many different exposure mappings $h(\vec{D}, i)$ 
\begin{equation}\label{eq:dgp_general}
    y_{it} = \mu_t + \mu_i + 2 D_{it} + \beta_{\text{spill, control}} (1-D_{it}) h(\vec{D}, i) + \varepsilon_{it}   
\end{equation}

@@

The group-time average direct effect is the building block of analysis in the framework proposed by C\&S. Estimation of the group-time average direct treatment effect, $ATT_{\text{direct}}(g,t)$ depends on which parallel trends assumption a researcher use. If a researcher is using Assumption \ref{eq:parallel_es_never}, then estimation will be done using a supsample containing only control units and units in group $g$. The difference-in-difference estimate uses period $t$ outcomes as the post-period and some period $g - \delta$ as the pre-period where $\delta$ is the number of periods before initial treatment. In the context of spillovers, the canoncial difference-in-differences estimation will have to be adjusted using a parameterized exposure mapping to prevent the bias detailed in the $2 \times 2$ setting. As shown in Section \ref{sec:monte_carlo}, this can be done using a set of donut indicators that is wide enough to capture all of the spillovers.

@@

>>>>

The rest of the paper is structured as follows. Section \ref{sec:po_framework} presents the potential outcomes framework, defines the estimand of interest, and shows the resulting bias from estimating a classical difference-in-differences model. Section \ref{sec:estimation} presents Monte Carlo simulations to illlustrate the bias result and to evaluate currently used solutions in the literature. Section \ref{sec:tva} presents an application for evaluating the Tennessee Valley Authority program. Section \ref{sec:event_study} discusses briefly how to incorporate spillovers into event-study estimation following the insights from \citet{Callaway_Sant’Anna_2020}.
% @TODO Fix this

@@

\begin{example}
    Last, is the spatial decay function where exposure decreases with distance. In this case, spillover intensity is the sum across all treated observations' decay term, i.e. \begin{equation}\label{eq:h_decay}
        h(\vec{D}, i) = \sum_{j \neq i} D_j e^{-\alpha d(i,j)}.
    \end{equation} 
    This exposure mapping allows for the intensity of spillovers to depend on distance to treatment and also is additive in the number of nearby units treated.\footnote{However, this specification assumes that all units are affected by all other units. This creates problems with inference because it implies potential correlation between all units' error terms. For this reason, this function often is summed over only the $k$-nearest neighbors or over units within $\bar{d}$ miles.} The speed of spillover decay over distance depends on the parameter $\alpha$ that can be calibrated by the researcher or estimated using non-linear least squares. In the literature on R\&D investment, for example, \citet{Keller_2002} uses a modified version of this exposure mapping where $D_j$ is country $j$'s R\&D expenditure. 
\end{example}

@@

\begin{proposition}[Decomposition of Difference-in-Differences Estimate]\label{thm:bias}\ \\    
    If Assumption \ref{eq:parallel} holds, the expectation of the estimate $\hat{\tau}$ from (\ref{eq:twfe}) is
    \begin{align*}
        \mathbb{E}[\hat{\tau}] &= \underbrace{\mathbb{E}\left[ Y_{i1} - Y_{i0} \ \vert \ D_i = 1 \right] - \mathbb{E}\left[ Y_{i1} - Y_{i0} \ \vert \ D_i = 0 \right]}_{\text{Difference-in-Differences}} \\ 
        &= 
        \mathbb{E} \left[ Y_{i1}(1, \vec{0}) - Y_{i1}(0, \vec{0}) \mid D_i = 1 \right] + \mathbb{E} \left[ Y_{i1}(1, h(\vec{D}, i)) - Y_{i1}(1, \vec{0}) \mid D_i = 1 \right] \\
        &\quad - \mathbb{E} \left[ Y_{i1}(0, h(\vec{D}, i)) - Y_{i1}(0, \vec{0}) \mid D_i = 0 \right] \\
        &= \tau_{\text{direct}} + \tau_{\text{spill,treated}} - \tau_{\text{spill,control}}
    \end{align*}
\end{proposition}

@@

% ------------------------------------------------------------------------------
\section{Estimation of Treatment Effects with Spillovers}
\label{sec:estimation}
% ------------------------------------------------------------------------------

In this section, I analyze methods that have been used by researchers for estimation of the direct and spillover effects. I provide practical guidance on when the methods are effective and in particular propose a method that is most robust.

% ------------------------------------------------------------------------------
\subsection{Removing Bias by Dropping Control Units}
% ------------------------------------------------------------------------------

A common approach in the literature is to drop control units neighboring treated units to avoid the problems of spillovers in identification. The above proposition would suggest that this can be effective in removing (at least part of) the average spillover effect on control units. However, since researchers do not know how far the spillovers spread there is difficulty in this method. On the one hand, removal of too few units will keep some control units that are experiencing spillover effects which would leave the treatment effect biased. On the other hand, removing too many units will leave few observations and therefore make the estimate have high variance. There is therefore a bias-variance trade off inherit to this method that can be readily improved by the method proposed in Section \ref{sec:parameterize}.

The second problem with this approach is that removal of control problems will not remove spillovers onto also treated units. Therefore if treated units are affected by the treatment status of nearby units, then the average spillover effect on treated untis which would remain as a source of bias. Appendix \ref{sec:monte_carlo} contains a set of Monte Carlo simulations that highlight the problems more rigorously.

% ------------------------------------------------------------------------------
\subsection{Parameterization of Spillovers}
\label{sec:parameterize}
% ------------------------------------------------------------------------------

While some researchers try to remove untis that are most likely to experience spillover effects in order to estimate the direct effects, another common approach is to directly parameterize the spillovers themselves. Parameterization of the potential outcomes allows for unbiased estimates of $\tau_{\text{direct}}$ by controlling for the exposure mapping directly. This is recommended for two reasons.

First, controlling for exposure mapping allows for unbiased estimates of the direct effect of treatment if the potential outcomes are correctly specified. However, the functional form of spillovers is not observed by the researcher. I show in simulations below that information on the exact functional form is not required for unbiased estimates of the treatment effect. Therefore, estimation can occur without dropping units which will improve efficiency.

Second, the spillover effects themselves are potentially relevant for policy makers. For example, if treatment for a given area is positive and there are positive benefits to neighboring regions, the estimated benefits will double-understate the true effect of treatment. First, it will not include the neighboring units' benefit in the calculation. Second, since the treatment effect will subtract the neighboring units' benefit, the neighboring units' benefit will be \emph{double undercounted}. Similarly, a policy maker trying to determine the net benefits for a treated location must consider the direct effect and the spillover effect onto treated units. This is because the size of benefits created by the treatment depend on the number of treated units nearby.

@@

Second, a researcher must decide how far the spillovers extend and whether the effect decays over distance. Things such as driving distance or access via public transportation can be helpful in deciding which units are `close enough' to experience spillovers. There are two commonly used techniques to model treatment effect decay.\footnote{If the spillovers is due to crossing borders into the treated area, then a reasonable decision about whether the effect decays is the frequency in which people use the good. For example, libraries whose benefit is only experienced from frequent usage would suggest that the effect decays over distance. On the other hand, abortion clinics opening are used infrequently and therefore the effect likely doesn't decay over distance within a certain cutoff.} First, as seen above is to use an exponential decay function with regards to distance. The parameter $\alpha$ in Equation (\ref{eq:h_decay}) determines the rate of decline over distance.

The other option is to use a set of `concentric rings' which are a semi-parametric way to estimate spillovers. If the spillovers do decay over distance, then the coefficients on the indicators should decrease as they move out. On the other hand, they are robust in estimating spillovers that do not decay, whereas exponential decay is not. In the simulations below, we will see that the rings are very flexible across forms of spillovers. An example of this is shown in the empirical application in Section \ref{sec:tva}.

Last, it is important for estimation to know if the spillovers are additive or non-additive in nature. If the number of nearby treated units matter, then the exposure mapping should sum over nearby treated units in some way. For example, in the context of library construction, the number of neighbors with libraries does not change the size of the spillover for it is access to \textit{any} library that matters. In the context of factory openings, though, agglomeration economies imply that the spillovers are additive in nature. The simulations below show that in order to accurately estimate spillovers requires correctly using a additive or non-additive version of the concentric rings.

To highlight the importance of these considerations, we turn to simulations where we include potentially incorrectly specified exposure mappings and see their performance in a set of different data generating processes. In the general form, I generate data with the same fixed effects and error term as before, but for many different exposure mappings $h(\vec{D}, i)$ 
\begin{equation}\label{eq:dgp_general}
    y_{it} = \mu_t + \mu_i + 2 D_{it} + \beta_{\text{spill, control}} (1-D_{it}) h(\vec{D}, i) + \varepsilon_{it}   
\end{equation}

@@

The group-time average direct effect is the building block of analysis in the framework proposed by C\&S. Estimation of the group-time average direct treatment effect, $ATT_{\text{direct}}(g,t)$ depends on which parallel trends assumption a researcher use. If a researcher is using Assumption \ref{eq:parallel_es_never}, then estimation will be done using a supsample containing only control units and units in group $g$. The difference-in-difference estimate uses period $t$ outcomes as the post-period and some period $g - \delta$ as the pre-period where $\delta$ is the number of periods before initial treatment. In the context of spillovers, the canoncial difference-in-differences estimation will have to be adjusted using a parameterized exposure mapping to prevent the bias detailed in the $2 \times 2$ setting. As shown in Section \ref{sec:estimation}, this can be done using a set of donut indicators that is wide enough to capture all of the spillovers.

@@

% ------------------------------------------------------------------------------
\section{Details on Monte Carlo Simulations}
\label{sec:monte_carlo}
% ------------------------------------------------------------------------------

I use the set of counties in the contiguous United States and the data generating process used is defined by the exposure mapping given in (\ref{eq:example_exposure}) with cutoff distance $\bar{d} = 40$ miles and the data generating process given in (\ref{eq:example_po}). A unit of observation is a US county and the periods are $t \in \{1, \dots, 20\}$ with treatment turning on after period $t = 10$. The unit and time fixed effects are generated by $\mu_t \sim N(0.2t, 0.1^2)$ and $\mu_i \sim N(6, 2^2)$ respectively, and the error term is $\varepsilon \sim N(0, 2^2)$. Last, the size of treatment and spillovers are as follows: $\beta_{\text{direct}} = 2$, $\beta_{\text{spill, control}} = 1$ and $\beta_{\text{spill, treat}} = 0$.

This data generating process where spillovers only occur onto control units matches what has been typically assumed in the literature. I assign treatment among counties randomly with various unconditional probabilities between 3 percent and 50 percent. The data-generating process is therefore 
\begin{equation}
    \label{eq:dgp1} 
    y_{it} = \mu_t + \mu_i + \beta_{\text{direct}} D_{it} + \beta_{\text{spill, control}} (1-D_{it}) \text{Near}_{it} + \varepsilon_{it}   
\end{equation}
